---
title: "Top Songs of the 1980's"
type: "generic_page"
---

<style>
p.caption {
 font-size: 10pt; 
}
</style>

# Top Songs of the 1980's

![Photo by [Israel Palacio](https://unsplash.com/@othentikisra?utm_source=unsplash&utm_medium=referral&utm_content=creditCopyText) on [Unsplash](https://unsplash.com/s/photos/radio?utm_source=unsplash&utm_medium=referral&utm_content=creditCopyText)](/./1980s_files/mic.jpg)


```{r setup, include=FALSE}
library(tidyverse)
library(slumdown)
library(knitr)
source('../data_prep.r')
source('../helper_functions.r')
page_year <- 1980

```

### Top 5 Songs
```{r top_5, echo=FALSE}
create_top_5(page_year)
```
### Most Popular Genres 
Here are the genre types that make up the top songs of the decade: 
<br>  
<br>
<center>
```{r top_genre, echo=FALSE}
create_top_genre(page_year)

```
</center>
<br> 
Rock music dominated the charts, with 'album rock' and 'dance rock' making up over half of the Top 100 songs of the decade. 'Adult standards', which is a genre defined by classic or oldschool hits of the past, also contributed to the top songs; followed by other subgenres of pop and rock. 
<br>
<br>


### What Makes a Song Popular 

Let's unpack what actually *made* a song popular in the `r page_year`'s. What elements of the music were important for it to be liked by the masses throughout a whole decade? To explore this, I ran a stepwise regression on all the audio features Spotify provided for the Top 100 songs of the `r page_year`'s, with popularity as the dependent variable. 

After running the analysis, we are left with the following co-efficients for a regression formula: 

<br> <center>
$$
popularity = 79.29 + (-.0.09)(energy) + (-0.02)(duration) + (-0.08)(acousticness) + (0.41)(speechiness)
$$
</center>
This equation lets us make the inference that in the `r page_year`'s, the audio features of a track that contributed to it being popular were energy, duration, acousticness and speechiness. More specifically, a song being lower in energy, shorter in duration, lower in acousticness and higher in speechiness would make it more popular.

A summary of the stepwise analysis is presented below:  
```{r stepwise, echo=FALSE} 
stepwise_popular_model(page_year)
```