---
title: "Top Songs of the 2000's"
type: "generic_page"
---

<style>
p.caption {
 font-size: 10pt;  
}
</style> 

# Top Songs of the 2000's

![Photo by [Victor Rodvang](https://unsplash.com/@rodograph?utm_source=unsplash&utm_medium=referral&utm_content=creditCopyText) on [Unsplash](https://unsplash.com/s/photos/pop-music?utm_source=unsplash&utm_medium=referral&utm_content=creditCopyText)](/./2000s_files/yellow_singer.jpg)


```{r setup, include=FALSE}
library(tidyverse)
library(slumdown)
library(knitr)
source('../data_prep.r')
source('../helper_functions.r')
page_year <- 2000

```

### Top 5 Songs
```{r top_5, echo=FALSE}
create_top_5(page_year)
```
### Most Popular Genres 

Here are the genre types that make up the top songs of the decade: 
<br>
<br>
<center>
```{r top_genre, echo=FALSE}
create_top_genre(page_year)
```
</center>
<br> 
Dance pop is the genre that defined the popular music of the 2000's. 'Barbadian pop' is the next most common genre; this ranking being led by Rihanna's emergence within and domination of the pop music scene. 

### What Makes a Song Popular 

Let's unpack what actually *made* a song popular in the `r page_year`'s. What elements of the music were important for it to be liked by the masses throughout a whole decade? To explore this, I ran a stepwise regression on all the audio features Spotify provided for the Top 100 songs of the `r page_year`'s, with popularity as the dependent variable. 

After running the analysis, we are left with the following co-efficients for a regression formula: 
<br>  
$$
popularity = 75.98 + (-.09)(energy)
$$
This equation lets us make the inference that in the `r page_year`'s, the audio features of a track that contributed to it being popular were energy only. More specifically, a song being lower in energy would make it more popular, although this is not a strong predictor.   

A summary of the stepwise analysis is presented below: 
```{r stepwise, echo=FALSE} 
stepwise_popular_model(page_year)
```