---
title: "Top Songs of the 2010's"
type: "generic_page"
---

<style> 
p.caption {
 font-size: 10pt;   
}
</style>       

# Top Songs of the 2010's

![Photo by [Austin Neill](https://unsplash.com/@arstyy?utm_source=unsplash&utm_medium=referral&utm_content=creditCopyText) on [Unsplash](https://unsplash.com/?utm_source=unsplash&utm_medium=referral&utm_content=creditCopyText)](/./2010s_files/smoke_dj.jpg)


```{r setup, include=FALSE}
library(tidyverse)
library(slumdown)
library(knitr)
source('../data_prep.r')
source('../helper_functions.r')
page_year <- 2010

```

### Top 5 Songs
```{r top_5, echo=FALSE}
create_top_5(page_year)   
```
### Most Popular Genres 

Here are the genre types that make up the top songs of the decade: 
<br>
<br>
<center>
```{r top_genre, echo=FALSE}      
create_top_genre(page_year)      

```
</center>
<br>     
With almost half of 2010's Top 100 songs fitting into the 'dance pop' genre, this represents what popular music sounded like in the decade. 'Big room' is a subgenre of electronic house music, and comes in second to the popularity of 'dance pop'.
<br>
<br>


### What Makes a Song Popular 

Let's unpack what actually *made* a song popular in the `r page_year`'s. What elements of the music were important for it to be liked by the masses throughout a whole decade? To explore this, I ran a stepwise regression on all the audio features Spotify provided for the Top 100 songs of the `r page_year`'s, with popularity as the dependent variable. 

After running the analysis, we are left with the following co-efficients for a regression formula: 
<br>  
$$
popularity = 89.05 + (-.21)(energy) + (0.17)(speechiness)
$$
This equation lets us make the inference that in the `r page_year`'s, the audio features of a track that contributed to it being popular were energy and speechiness. More specifically, a song being lower in energy and higher in speechiness would make it more popular.      

A summary of the stepwise analysis is presented below:  
```{r stepwise, echo=FALSE} 
stepwise_popular_model(page_year)
```